---
title: "**Computer Intensive Methods - Final projects (2022)**"
output: 
  bookdown::pdf_document2:
      latex_engine: xelatex
      toc: true
      toc_depth: 3
      number_sections: true
      # df_print: kable
      highlight: kate 
subtitle:  |
  **Group **: 
  Deo Byabazaire (2159254)  
  Mirriam Dianah Lucheveleli (2159277)  
  Farida Iddy (2159270)  
  Quynh Long Khuong (2159280)  
    
fontsize: 11pt
mainfont: Times New Roman #Georgia
geometry: margin=1in
layout: page
keep-latex: true
links-as-notes: true
urlcolor: orange
linkcolor: orange
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,	
                      warning = FALSE,
                      eval = TRUE,
                      tidy = FALSE)

library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
```


\newpage



# Project 1


```{r}
library("DAAG")
data(nassCDS)
names(nassCDS)
dim(nassCDS)
# Check missing value
sapply(nassCDS, function(x){sum(is.na(x))})
# complete-case data
nassCDS <- na.omit(nassCDS) %>% mutate(dead = ifelse(dead == "dead", 1, 0))
dim(nassCDS)
```

## Question 1

Let $Y_i$ be an indicator variable which takes the value of 1 if an occupant died in an accident (the variable `dead`) and zero otherwise and $X_i$ be the age of occupant in years (the variable `ageOFocc`). We consider the following GLM

$$g(P(Y_i = 1)) = \beta_0 + \beta_1X_i$$

1. Estimate the model using the classical GLM approach


```{r}
glm_dead <- glm(dead ~ ageOFocc, data = nassCDS, family = "binomial")
summary(glm_dead)
```
2. Let $X_50$ be the age of occupant for which the probability to die is 0.5 $P(Y_i = 1) = 0.5$. Estimate $X_50$. Use non parametric bootstrap to estimate the distribution of $X_50$
and construct a 95% for the $X_50$

```{r}
B <- 1000
n <- length(nassCDS$dead)
index <- c(1:n)
X_50 <- c()

for (i in seq(B)) {
    index.b <- sample(index, n, replace=TRUE)
    dead <- nassCDS$dead[index.b]
    ageOFocc <- nassCDS$ageOFocc[index.b]
    glm_dead2 <- glm(dead ~ ageOFocc, family = "binomial")
    X_50[i] <- (-coef(glm_dead2)[[1]])/coef(glm_dead2)[[2]]
}
# Distribution of X50
hist(X_50)
# Estimate 95% CI
c(quantile(X_50, 0.025), quantile(X_50, 0.975))
```
3. For the model formulated above, we focus on the odds ratio (OR) for a unit increased
in age. Estimate the OR and Use parametric bootstrap to test the null hypothesis
$H_0: OR = 1$

```{r}
glm_dead_null <- glm(dead ~ 1, data = nassCDS, family = "binomial")
summary(glm_dead_null)
```

```{r}
glm_dead_null <- glm(dead ~ 1, data = nassCDS, family = "binomial")
eta <- coef(glm_dead_null)[[1]]
prob_dead <- exp(eta)/(1 + exp(eta))

B <- 1000
n <- length(nassCDS$dead)

beta0 <- c()
beta1 <- c()

for (i in seq(B)) {
    dead_pred <- rbinom(n, 1, prob_dead)
    glm_dead2 <- glm(dead_pred ~ nassCDS$ageOFocc, family = "binomial")
    beta0[i] <- coef(glm_dead2)[[1]]
    beta1[i] <- coef(glm_dead2)[[2]]
}
```

```{r}
beta1_hat <- coef(glm_dead)[[2]] # Observed beta1
beta1 %>% as_tibble() %>% 
    ggplot(aes(x = beta1)) + 
    geom_histogram(fill = "navy", color = "white") + 
    geom_vline(xintercept = beta1_hat, color = "red") +
    theme_minimal()


pvalue <- (1 + sum(abs(beta1_hat) < abs(beta1))) /(B+1)
pvalue
```


4. Let $\pi_{33}$ be the probability of death for an occupant at age 33. Use parametric
bootstrap to calculate the standard error for $\hat{\pi_{33}}$ and construct a 90% bootstrap t C.I. for $\pi_{33}$


```{r}
beta0_hat <- coef(glm_dead)[[1]] 
beta1_hat <- coef(glm_dead)[[2]] 

eta_hat <- beta0_hat + beta1_hat*33
prob_dead_hat <- exp(eta_hat)/(1 + exp(eta_hat))

B <- 1000
n <- length(nassCDS$dead)

prob_dead332 <- c()

for (i in seq(B)) {
    dead_pred <- rbinom(n, 1, prob_dead_hat)
    glm_dead2 <- glm(dead_pred ~ 1, family = "binomial")
    beta0 <- coef(glm_dead2)[[1]]
    eta <- beta0
    prob_dead332[i] <- exp(eta)/(1 + exp(eta))
}

hist(prob_dead33)
sd(prob_dead33) # SE
# Estimate 95% CI
c(quantile(prob_dead33, 0.05), quantile(prob_dead33, 0.95))
```

## Question 2

In this question we use the same model formulated in Q1. Let the parameters vector be
$\eta = (\beta_0, \beta_1)$ and the parameter estimates be $\hat{\eta} = (\hat{\beta_0}, \hat{\beta_1})$. Use parametric and non parametric bootstrap in order to:

1. Approximate the distribution of $\hat{\eta}$, i.e., the joint distribution of $\hat{\beta_0}$ and $\hat{\beta_1}$.
2. Construct a joint 95% confidence interval for $\eta$

### Non-parametric bootstrap


```{r}
B <- 1000
n <- length(nassCDS$dead)
index <- c(1:n)
beta0 <- c()
beta1 <- c()

for (i in seq(B)) {
    index.b <- sample(index, n, replace=TRUE)
    dead <- nassCDS$dead[index.b]
    ageOFocc <- nassCDS$ageOFocc[index.b]
    glm_dead2 <- glm(dead ~ ageOFocc, family = "binomial")
    beta0[i] <- coef(glm_dead2)[[1]]
    beta1[i] <- coef(glm_dead2)[[2]]
}
# Joint distribution
eta = c(beta0, beta1)
hist(eta)
# Estimate 95% CI
c(quantile(eta, 0.05), quantile(eta, 0.95))
```

### Parametric bootstrap

```{r}
beta0_hat <- coef(glm_dead)[[1]]
beta1_hat <- coef(glm_dead)[[2]]


B <- 1000
n <- length(nassCDS$dead)

beta0 <- c()
beta1 <- c()
dead <- c()

for (i in seq(B)) {
    for (j in seq(n)) {
        eta <- beta0_hat + beta1_hat*nassCDS$ageOFocc[j]
        yprob <- exp(eta)/(1 + exp(eta))
        dead[j] <- rbinom(1, 1, yprob)
    }
    glm_dead2 <- glm(dead ~ nassCDS$ageOFocc, family = "binomial")
    beta0[i] <- coef(glm_dead2)[[1]]
    beta1[i] <- coef(glm_dead2)[[2]]
}

# Joint distribution
eta = c(beta0, beta1)
hist(eta)
# Estkimate 95% CI
c(quantile(eta, 0.05), quantile(eta, 0.95))
```


## Question 3

In this question we focus of the following 2 × 2 table (for the complete case analysis) for the
variables airbag and dead


```{r}
nassCDS %<>% mutate(airbag = ifelse(airbag == "airbag", 1, 0))
```


1. Define the observation unit $(X_i, Y_i)$ for the question


```{r}
table(nassCDS$airbag, nassCDS$dead)
```



2. Calculate the odds ratio for usage of airbag and the accident outcome (dead/alive)
and construct 95% confidence interval. You can use the R function `oddsratio`. What is your conclusions? Do you think that airbags in the car influence the accident outcome ?


```{r}
library(epitools)
x <- oddsratio(nassCDS$airbag, nassCDS$dead)$measure
x[2, ][[1]] # Odds ratio
c(x[2, ][[2]], x[2, ][[3]])

```

3. Use parametric bootstrap to construct a construct a 95% confidence interval for the OR


```{r}
nn <- length(nassCDS$dead)
m <- 1180
n <- 24883
k <- 14336
a <- rhyper(nn, m, n, k)
mean(a)
```

```{r}
b <- k - a
c <- m - a
d <- n - k - a
OR <- a*d/(b*c)
hist(OR)
```


```{r}

```


```{r}

```



```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


